### 2012-11-21 paper.tex
worked on decision trees for echocardiogramm

\usepackage{natbib}
\bibliographystyle{plain}

@@

\begin{verbatim}
J48 pruned tree
------------------

survival <= 12
|   survival <= 11: 0 (4.0)
|   survival > 11: 1 (4.0)
survival > 12: 1 (87.0)
\end{verbatim}

@@

The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute don't make any sense ($ 11 $ is yet a smaller number than $ 12 $, isn't it?). E.g. choosing a DecisionStump results in the expected single splitting: 
\begin{verbatim}
Class distributions
survival <= 11.5
0   1   
1.0 0.0 
survival > 11.5
0   1   
0.0 1.0 
survival is missing
0   1   
0.042105263157894736    0.9578947368421052
\end{verbatim}
Since there is no missing value for the survival attribute, this branch can be ignored.

@@

\subsection{re-pre-processed the data set}

As mentioned in the description of the dataset on the UCI repo site we can remove the attributes group, name and mold as they do not have any meaning for the prediction attribute.

We also have to remove the survival and still-alive attributes, as the prediction attribute derives from them. Otherwise, a classification algorithm could predict all instances by taking into account only these two attributes.

@@

\section{Adult data}

>>>>

\usepackage{natbib}
\usepackage{listings}
\bibliographystyle{plain}

@@

\begin{lstlisting}
J48 pruned tree
------------------

survival <= 12
|   survival <= 11: 0 (4.0)
|   survival > 11: 1 (4.0)
survival > 12: 1 (87.0)
\end{lstlisting}

@@

The only remarkable thing is, that the J48 implementation under default settings seems to enforce a certain depth of the tree, even if the splitting on the same attribute don't make any sense ($ 11 $ is yet a smaller number than $ 12 $, isn't it?). E.g. choosing a DecisionStump results in the expected single splitting: 
\begin{lstlisting}
Class distributions
survival <= 11.5
0   1   
1.0 0.0 
survival > 11.5
0   1   
0.0 1.0 
survival is missing
0   1   
0.042105263157894736    0.9578947368421052
\end{lstlisting}
Since there is no missing value for the survival attribute, this branch can be ignored.

@@

\subsection{re-preprocessed the data set}

As mentioned in the description of the dataset on the UCI repository site we can remove the attributes ``group'', ``name'' and ``mult'' as they do not have any meaning for the prediction attribute.

We also have to remove the ``survival'' and ``still-alive'' attributes, as the prediction attribute derives from them. Otherwise, a classification algorithm could predict all instances by taking into account only these two attributes.

@@

\subsection{Decision Tree on re-preprocessed data}

The J48 Decision Tree shows a more useful tree on the re-preprocessed data:
\begin{lstlisting}
J48 pruned tree
------------------

epss <= 13100: 1 (90.63/1.95)
epss > 13100
|   fractional-shortening <= 190: 1 (2.25/0.05)
|   fractional-shortening > 190: 0 (2.11/0.11)
\end{lstlisting}

The documentation of the data set explains the two attributes as follows:
\begin{quote}
``epss -- E-point septal separation, another measure of contractility. Larger numbers are increasingly abnormal.''
\end{quote}
\begin{quote}
`` fractional-shortening -- a measure of contracility around the heart lower numbers are increasingly abnormal''
\end{quote}

Only the rule for the fractional shortening should be the other way round, because a low fractional shortening is a rather pathological sign. Although the decision tree has a high precision of almost 95\% it performs very imprecise, because none of the patients who died after one year are correctly classified:

\begin{lstlisting}
=== Confusion Matrix ===

  a  b   <-- classified as
  0  4 |  a = 0
  1 90 |  b = 1

\end{lstlisting}

Using an unpruned tree doesn't improve the results:
\begin{lstlisting}
J48 unpruned tree
------------------

pericardial-effusion = 0
|   epss <= 13100: 1 (76.62/1.95)
|   epss > 13100
|   |   epss <= 15600: 0 (2.19/0.16)
|   |   epss > 15600: 1 (2.19/0.03)
pericardial-effusion = 1: 1 (14.0)
\end{lstlisting}
Here, the existence of a pericardial effusion (fluid around the heart) is assumed to be a healthy characteristic, but it is not. Accordingly, the precision of the J48 decreases from 95\% to 88\%.
\section{Adult data}
